\subsection{ Graph Coloring Benchmark } \label{sec:graph-coloring-benchmark}

The graph coloring benchmark employs a graph coloring algorithm designed for distributed WLAN channel selection \citep{leith2012wlan}.
In this algorithm, nodes begin by randomly choosing a color.
Each computational update, nodes test for any neighbor with the same color.
If and only if a conflicting neighbor is detected, nodes randomly select another color.
The probability of selecting each possible color is stored in an array associated with each node.
Before selecting a new color, the stored probability of selecting the current (conflicting) color is decreased by a multiplicative factor $1 - b$.
We used $b=0.1$, as suggested by Leith et al.
Tandem adjustments increase the probability of selecting all other channels.
Regardless of whether their color changed, nodes always transmit their current color to their neighbors.

Our benchmarks focus on weak scalability, using a fixed problem size of \numprint{2048} graph nodes per thread or process.
These nodes were arranged in a two-dimensional grid topology where each node had three possible colors and four neighbors.
We implement the algorithm with a single Conduit communication layer carrying graph color as an unsigned integer.
We used Conduit's built-in pooling feature to consolidate color information into a single MPI message between pairs of communicating processes each update.
We performed five replicates, each with a five second simulation runtime.
Solution error was measured as the number of graph color conflicts remaining at the end of the benchmark.
