\subsection{Quality of Service: Intranode vs. Internode}
\label{sec:intranode-vs-internode}

\input{fig/intranode-vs-internode/intranode-vs-internode.tex}

This section tests the effect of process assignment on best-effort quality of service, comparing multi-node and single-node assignments.
The graph coloring benchmark again serves as our experimental substrate.

For this experiment, processes were either assigned to the same node or were assigned to different nodes.
In both cases, we used two processes.

\subsubsection{Simstep Period}

Simstep period was significantly slower under internode conditions than under intranode conditions.

% When processes shared the same node, simstep period was around \SI{9}{\micro\second} (mean: \SI{9.07}{\micro\second} inlet / \SI{9.04}{\micro\second} outlet; median: \SI{9.11}{\micro\second} inlet / \SI{9.05}{\micro\second}).
When processes shared the same node, simstep period was around \SI{9}{\micro\second} (mean: \SI{9.06}{\micro\second}; median: \SI{9.08}{\micro\second}).
% Under internode conditions, simstep period was around \SI{14}{\micro\second} (mean: \SI{14.5}{\micro\second} inlet/outlet; median: \SI{14.5}{\micro\second} inlet / \SI{14.4}{\micro\second} outlet).
Under internode conditions, simstep period was around \SI{14}{\micro\second} (mean: \SI{14.5}{\micro\second}; median: \SI{14.4}{\micro\second}).
Supplementary Figures \ref{fig:intranode-vs-internode-distribution-simstep-period-inlet-ns} and \ref{fig:intranode-vs-internode-distribution-simstep-period-outlet-ns} depict the distribution of walltime per computational update across intranode and internode conditions.

This result presumably attributes to an increased walltime cost for calls to the MPI implementation backing internode communication compared to the MPI implementation backing intranode communication.
Although this effect is clearly detectable, its magnitude is modest given the minimal computational intensity of the simulation update step --- only $\approx 56\%$ more expensive than intranode dispatch.

Both mean and median simstep period increased significantly under internode conditions.
(Supplementary Figures \ref{fig:intranode-vs-internode-regression-simstep-period-inlet-ns} and \ref{fig:intranode-vs-internode-regression-simstep-period-outlet-ns} visualize these regressions and Supplementary Tables \ref{tab:intranode-vs-internode-ordinary-least-squares-regression} and \ref{tab:intranode-vs-internode-quantile-regression} detail numerical results.)

\subsubsection{Simstep Latency}

Significantly more simulation updates transpired during message transmission under internode conditions compared to intranode conditions.

Supplementary Figures \ref{fig:intranode-vs-internode-distribution-latency-simsteps-inlet} and \ref{fig:intranode-vs-internode-distribution-latency-simsteps-outlet} compare the distributions of simstep latency across these conditions.
% Simstep latency was around 1 update for intranode communication (mean: 1.01 updates inlet / 0.99 updates outlet; median 0.75 updates inlet / outlet) and around 40 updates for internode communication (mean: 41.8 updates inlet / 41.4 updates outlet; median: 37.4 updates inlet / outlet).
Simstep latency was around 1 update for intranode communication (mean: 1.00 updates; median 0.75 updates) and around 40 updates for internode communication (mean: 41.6 updates; median: 37.4 updates).

Regression analysis confirms the significant effect of process placement on simstep latency (Supplementary Figures \ref{fig:intranode-vs-internode-regression-latency-simsteps-inlet} and \ref{fig:intranode-vs-internode-regression-latency-simsteps-outlet}).
Supplementary Tables \ref{tab:intranode-vs-internode-ordinary-least-squares-regression} and \ref{tab:intranode-vs-internode-quantile-regression} detail numerical results of these regressions.

\subsubsection{Walltime Latency}

Significantly more walltime elapsed during message transmission under internode conditions compared to intranode conditions.

% Walltime latency was less than \SI{10}{\micro\second} for intranode communication (mean: \SI{7.72}{\micro\second} inlet / \SI{7.69}{\micro\second}; median: \SI{6.95}{\micro\second} inlet / \SI{6.93}{\micro\second}).
Walltime latency was less than \SI{10}{\micro\second} for intranode communication (mean: \SI{7.70}{\micro\second}; median: \SI{6.94}{\micro\second}).
% Internode communication had approximately $50\times$ greater walltime latency, at around \SI{500}{\micro\second} (mean: \SI{604}{\micro\second} inlet / \SI{596}{\micro\second} outlet; median: \SI{554}{\micro\second} inlet / \SI{548}{\micro\second} outlet).
Internode communication had approximately $50\times$ greater walltime latency, at around \SI{500}{\micro\second} (mean: \SI{600}{\micro\second}; median: \SI{551}{\micro\second}).

Supplementary Figures \ref{fig:intranode-vs-internode-distribution-latency-walltime-inlet-ns} and \ref{fig:intranode-vs-internode-distribution-latency-walltime-outlet-ns} show the distributions of walltime latency for intra- and inter-node communication.
Regression analysis confirmed a significant increase in walltime latency under inter-node communication (Supplementary Figures \ref{fig:intranode-vs-internode-regression-latency-walltime-inlet-ns}, \ref{fig:intranode-vs-internode-regression-latency-walltime-outlet-ns}; Supplementary Tables \ref{tab:intranode-vs-internode-ordinary-least-squares-regression} and \ref{tab:intranode-vs-internode-quantile-regression}).

\subsubsection{Delivery Clumpiness}

Delivery clumpiness was minimal under intranode communication and very high under internode communication.

Under intranode conditions, we observed a mean clumpiness value of 0.014 and a median of 0.002.
Under internode conditions, we observed mean and median clumpiness values of 0.96.
Supplementary Figures \ref{fig:intranode-vs-internode-distribution-delivery-clumpiness} and \ref{fig:intranode-vs-internode-distribution-delivery-clumpiness} show the distributions of clumpiness for intra- and inter-node communication.
Regression analysis confirmed a significant increase in clumpiness under inter-node communication (Supplementary Figures \ref{fig:intranode-vs-internode-regression-delivery-clumpiness}, \ref{fig:intranode-vs-internode-regression-delivery-clumpiness}; Supplementary Tables \ref{tab:intranode-vs-internode-ordinary-least-squares-regression} and \ref{tab:intranode-vs-internode-quantile-regression}).

\subsubsection{Delivery Failure Rate}

Somewhat counterintuitively, a significantly higher proportion of deliveries failed for intranode communication than for internode communication.

We observed a delivery failure rate of around 0.3 for intranode communication (mean: 0.33; median: 0.30) and no delivery failures for internode communication (mean: 0.00; median: 0.00).
In some intranode snapshot windows, we observed a delivery failure rate as high as 0.8.
Supplementary Figures \ref{fig:intranode-vs-internode-distribution-delivery-clumpiness} and \ref{fig:intranode-vs-internode-distribution-delivery-clumpiness} show the distributions of delivery failure rate for intra- and inter-node communication.

Because of Conduit's current MPI-based implementation, messages only drop when the underlying send buffer fills; queued messages are guaranteed for delivery.
Slower simstep period under internode allocation could improve stability of the send buffer due to more time, on average, between send attempts.
Underlying buffering or consolidation by the MPI backend for internode communication might also play a role by allowing data to be moved out of the userspace send buffer more promptly.

Regression analysis confirmed a significant increase in delivery failure under intra-node communication (Supplementary Figures \ref{fig:intranode-vs-internode-regression-delivery-clumpiness}, \ref{fig:intranode-vs-internode-regression-delivery-clumpiness}; Supplementary Tables \ref{tab:intranode-vs-internode-ordinary-least-squares-regression} and \ref{tab:intranode-vs-internode-quantile-regression}).
