\subsection{Quality of Service: Faulty Hardware}
\label{sec:with-lac-417-vs-sans-lac-417}

\input{fig/with-lac-417-vs-sans-lac-417/with-lac-417-vs-sans-lac-417.tex}

The extreme magnitude of outliers for metrics reported in Section \ref{sec:weak-scaling} prompted further investigation of the conditions under which these outliers arose.
Closer inspection revealed that the most extreme outliers were all associated with snapshots on a single node: lac-417.

So, we acquired two separate 256 process allocations on the lac cluster: one including lac-417 and one excluding lac-417.

Supplementary Figures \ref{fig:with-lac-417-vs-sans-lac-417-distribution-simstep-period-inlet-ns}, \ref{fig:with-lac-417-vs-sans-lac-417-distribution-simstep-period-outlet-ns}, \ref{fig:with-lac-417-vs-sans-lac-417-distribution-latency-simsteps-inlet}, \ref{fig:with-lac-417-vs-sans-lac-417-distribution-latency-simsteps-outlet}, \ref{fig:with-lac-417-vs-sans-lac-417-distribution-latency-walltime-inlet-ns}, \ref{fig:with-lac-417-vs-sans-lac-417-distribution-latency-walltime-outlet-ns}, \ref{fig:with-lac-417-vs-sans-lac-417-distribution-delivery-clumpiness}, and \ref{fig:with-lac-417-vs-sans-lac-417-distribution-delivery-failure-rate} compare the distributions of quality of service metrics between allocations with and without lac-417.
Extreme outliers are present exclusively in the lac-417 allocation for walltime latency, simstep latency, and delivery failure rate.
Otherwise, the metrics' distributions across snapshots are very similar between allocations.

Supplementary Figures \ref{fig:with-lac-417-vs-sans-lac-417-regression-simstep-period-inlet-ns}, \ref{fig:with-lac-417-vs-sans-lac-417-regression-simstep-period-outlet-ns}, \ref{fig:with-lac-417-vs-sans-lac-417-regression-latency-simsteps-inlet}, \ref{fig:with-lac-417-vs-sans-lac-417-regression-latency-simsteps-outlet}, \ref{fig:with-lac-417-vs-sans-lac-417-regression-latency-walltime-inlet-ns}, \ref{fig:with-lac-417-vs-sans-lac-417-regression-latency-walltime-outlet-ns}, \ref{fig:with-lac-417-vs-sans-lac-417-regression-delivery-clumpiness}, \ref{fig:with-lac-417-vs-sans-lac-417-regression-delivery-clumpiness}, and
\ref{fig:with-lac-417-vs-sans-lac-417-regression-delivery-failure-rate} chart OLS and quantile regressions of quality of service metrics on job composition.
Mean walltime latency, simstep latency, and delivery failure rate are all significantly greater with lac-417.
Surprisingly, mean simstep period is significantly longer without lac-417.

However, there is no significant difference in median value for any quality of service metric between allocations including or excluding lac-417.
This stability of metric medians within allocations containing lac-417 --- which have significantly different means due to outlier values induced by the presence of lac-417 --- demonstrates how the best-effort system maintains overall quality of service stability despite defective or degraded components.

Supplementary Tables \ref{tab:with-lac-417-vs-sans-lac-417-ordinary-least-squares-regression} and \ref{tab:with-lac-417-vs-sans-lac-417-quantile-regression} provide numerical details on regressions reported above.
