\begin{abstract}
% Developing software to effectively take advantage of growth in parallel and distributed processing capacity poses significant challenges.
% Best-effort computing models have been proposed as a strategy to overcome challenges around harnessing high performance computing at extreme scale.
% These models relax synchronization and hardware reliability requirements, accepting nondeterminism in exchange for efficiency.
% This trade-off, inherent to best-effort computing, is often compatible with nature-inspired computing methods.
% Although many programming languages and frameworks aim to facilitate software development for high performance applications, existing tools do not directly provide a prepackaged best-effort interface.
% The Conduit C++ Library aims to provide a convenient interface for best-effort inter-thread and inter-process communication.
% Here, we describe the motivation, objectives, design, and implementation of the library.
% Benchmarks on a communication-intensive graph coloring problem and a compute-intensive digital evolution simulation, show that Conduit's best-effort model can improve scaling efficiency and solution quality, particularly in a distributed, multi-node context.

% Conduit will be useful facilitated efficient, scalable parallel and distributed implementations of natural computing algorithms.

Developing software to effectively take advantage of growth in parallel and distributed processing capacity poses significant challenges.
Traditional programming techniques allow a user to assume that execution, message passing, and memory are always kept synchronized.
However, maintaining this consistency becomes increasingly costly at scale. 
One proposed strategy is ``best-effort computing'', which relaxes synchronization and hardware reliability requirements, accepting nondeterminism in exchange for efficiency.
Although many programming languages and frameworks aim to facilitate software development for high performance applications, existing tools do not directly provide a prepackaged best-effort interface.
The Conduit C++ Library aims to provide such an interface for convenient implementation of software that uses best-effort inter-thread and inter-process communication.
Here, we describe the motivation, objectives, design, and implementation of the library.
Benchmarks on a communication-intensive graph coloring problem and a compute-intensive digital evolution simulation show that Conduit's best-effort model can improve scaling efficiency and solution quality, particularly in a distributed, multi-node context.
%% Conduit will be useful facilitated efficient, scalable parallel and distributed implementations of natural computing algorithms. (this is why you care as a GP person)

\end{abstract}


